---
title: "Introduction to Strings/Text"
author: "STA 325, Supplemental Material"
date: Rebecca C. Steorts
output: 
     beamer_presentation:
      includes: 
          in_header: custom2.tex
font-size: 8px
---

Agenda
===
- Introduction to strings and string operations
- Extracting and manipulating string objects
- Introduction to general search

Why Characters?
================

Most data we deal with is in character form!

- web pages can be scraped
- emails can be analyzed for network properties
- survey responses must be processed and compared

Even if you only care about numbers, it helps to be able to extract them from text and manipulate them easily.


Characters versus Strings
========================

- ***Character***: a symbol in a written language, specifically what you can enter at a keyboard: letters, numerals, punctuation, space, newlines, etc.

```
'L', 'i', 'n', 'c', 'o', 'l'
```

- ***String***: a sequence of characters bound together

```
Lincoln
```

Characters/Strings
====

R does not have a separate type for characters and strings

```{r}
mode("L")
mode("Lincoln")
class("Lincoln")
```

Making Strings
==============

- To construct a string use single or double quotations.
- Use `nchar()` to get the length of a single string. 

Constructing Strings
====
\footnotesize
```{r}
"Lincoln"
"Abraham Lincoln"
"Abraham Lincoln's Hat"
"As Lincoln never said, \"Four score and seven beers ago\""
```




Whitespace
==========

- The space, `" "` is a character 

- Multiple spaces `"   "` and the empty string `""` are also characters

- Some characters are special, so we have "escape characters" to specify them in strings:

1. quotes within strings: `\"`
2. tab: `\t`
3. new line `\n` and carriage return `\r`  (use the former rather than the latter when possible)


The character data type
=========================

- The character data type is one of the atomic data types, like `numeric` or `logical`

- The character data type can go into scalars, vectors, arrays, lists, or can be the type of a column in a data frame. 

Constructing Strings 
===
```{r}
length("Abraham Lincoln's beard")
length(c("Abraham", "Lincoln's", "beard"))
nchar("Abraham Lincoln's beard")
nchar(c("Abraham", "Lincoln's", "beard"))
```

Character-Valued Variables
==========================

They work just like others, e.g., with vectors:

```{r}
president <- "Lincoln" # variable assignment
nchar(president)  # not 9
presidents <-
  c("Fillmore","Pierce",
    "Buchanan","Davis","Johnson")
presidents[3] # shows element 3
presidents[-(1:3)] # removes elements 1 through 3
```

Displaying Characters
=====================

- `print()`: directly prints to the console

- `cat()` writes the string directly to the console (removes the quotations)

- `message()`: Is for debugging purposes (read about more on your own from the R help files). 

Displaying Characters
=====================

```{r}
print("Abraham Lincoln")
cat("Abraham Lincoln")
cat(presidents)
message(presidents)
```


Substring Operations
================

***Substring***: This is a subset of a larger string. It's also a string in it's own right.  

A string is not a vector or a list, so we ***cannot*** use subscripts like `[[ ]]` or `[ ]` to extract substrings. 

To work with substring, we use  `substr()` instead.

We take the string and break it into a small string (substring) beginning with the start character and ending with the stop character. 

Substring Operations
================
\footnotesize
```{r}
phrase <- "Christmas Bonus"
substr (phrase, start=8, stop=12) 
```

We can also use `substr` to replace elements:

```{r}
substr(phrase, 13, 13) <- "g"
phrase
```

Does white space count as a character? Yes! 

```{r}
substr(phrase, 10, 10) <- "Z"
phrase
```


substr() for String Vectors
============================

`substr()` vectorizes over all its arguments.

Let's recall our variable that we created earlier of a few presidents.

\footnotesize
```{r}
presidents
```

substr() for String Vectors
============================
\footnotesize
```{r}
presidents
substr(presidents,1,2)   # First two characters
```

substr() for String Vectors
============================
\footnotesize

```{r}
presidents
nchar(presidents)
# Last two characters
substr(presidents,nchar(presidents)-1,nchar(presidents))   

```

substr() for String Vectors
============================
 

\footnotesize
```{r}
presidents
# No such substrings so return the null string
substr(presidents,20,21)    
```

substr() for String Vectors
============================



\footnotesize
```{r}
presidents
# Explain!
substr(presidents,7,7)      
```

Dividing Strings into Vectors
=============================

`strsplit()` divides a string according to key characters by splitting each element of the character vector `x` at appearances of the pattern `split`.

The output is a list of character vectors. 

Dividing Strings into Vectors
=============================
```{r}
scarborough.fair <- "parsley, sage, rosemary, thyme"
strsplit (scarborough.fair, ",")
strsplit (scarborough.fair, ", ")
```

Dividing Strings into Vectors
=============================

Pattern is recycled over elements of the input vector:
\footnotesize
```{r}
strsplit(c(scarborough.fair, 
           "Garfunkel, Oates", "Clement, McKenzie"), ", ")
```


Combining Vectors into Strings
==============================

Converting one variable type to another is called ***casting***: 

```{r}
as.character(7.2)            # Obvious
as.character(7.2e12)         # Obvious
as.character(c(7.2,7.2e12))  # Obvious
as.character(7.2e5)          # Not quite so obvious
```


Building strings from multiple parts
====================

The `paste()` function is very flexible!

With one vector argument, works like `as.character()`:

```{r}
paste(41:45)
```


Building strings from multiple parts
====================

- We can build strings from multiple parts easily. 

- For example, with 2 or more vector arguments,
we can combine them with recycling:

\scriptsize
```{r}
paste(presidents,41:45) # None are historically accurate!
paste(presidents,c("R","D"))
paste(presidents,"(",c("R","D"),41:45,")") 
```


A More Complicated Example of Recycling
=======================================

Exercise: Convince yourself of why this works as it does

\scriptsize
```{r}
rep(1:11,times=rep(2,11)) 
paste(c("HW","Lab"),rep(1:11,times=rep(2,11)))
```


Condensing Multiple Strings
===========================

Default value of `collapse` is `NULL` (meaning the function doesn't use it) 

Let's produce one large string and utilize `collapse`

\scriptsize
```{r}
paste(presidents, " (", 41:45, ")", sep="", collapse="; ")
```


Text of Some Importance
=========


  If we shall suppose that American slavery is one of those
  offenses which, in the providence of God, must needs come, but which, having
  continued through His appointed time, He now wills to remove, and that He
  gives to both North and South this terrible war as the woe due to those by
  whom the offense came, shall we discern therein any departure from those
  divine attributes which the believers in a living God always ascribe to Him?
  Fondly do we hope, fervently do we pray, that this mighty scourge of war may
  speedily pass away.  Yet, if God wills that it continue until all the wealth
  piled by the bondsman's two hundred and fifty years of unrequited toil shall
  be sunk, and until every drop of blood drawn with the lash shall be paid by
  another drawn with the sword, as was said three thousand years ago, so still
  it must be said "the judgments of the Lord are true and righteous
  altogether."
  
- Abraham Lincoln, Second Inaugural Address, 1865

Lincoln's Second Inaugural Address (1865)
=========
The full text can be found in `lincoln.txt`

Let's start by looking at a summary of this textual data 

\tiny
```{r}
al2 <- readLines("lincoln.txt")
length(al2)
head(al2)
```

`al2` is a vector, one element per line of text


A Hint Of The Future: Search
============================

Narrowing down entries: use `grep()` to find which strings have a matching search term

1. grep: returns the index value of each matched pattern

2. grepl: returns a logical output for each indices in the original vector with "TRUE" representing matched patterns

grep versus grepl
===

\scriptsize
```{r}
grep("God", al2)  
grepl("God", al2) 
al2[grep("God", al2)] # Returns the strings of matches
```

Exercises
===

There are some simple exercises (tasks) to test you knowledge
of this material. Please go through this on your own. 


Task 1: Breaking words apart
==============

Task 1: How would one make a long string using the text and then split all the words? 

Task 2:  Then how would we count how often each word appears and put these in order?

Task 3: What do we find that is unexpected and expected from this simple exercise? 

\scriptsize
```{r}
al2 <- paste(al2, collapse=" ")
al2.words <- strsplit(al2, split=" ")[[1]]
head(al2.words)
```

Task 2: Counting Words with table()
=============================

Tabulate how often each word appears, put in order:

\scriptsize
```{r}
wc <- table(al2.words)
wc <- sort(wc,decreasing=TRUE)
head(wc,20)
```

`names(wc)` gives all the distinct words in `al2.words` (***types***); `wc` counts how often they appear (***tokens***)




Task 3: Unexpected
==========

The null string is the third-most-common word:
```{r}
names(wc)[3]
```

```{r}
wc["years"]
wc["years,"]
```

Task 3: Unexpected
==========

Capitalization:
```{r}
wc["that"]
wc["That"]
```

All of this can be fixed if we learn how to work with text patterns and not just constants.

Summary
=======

- Text is data, just like everything else
- `substr()` extracts and substitutes
- `strsplit()` turns strings into vectors
- `paste()` turns vectors into strings
- `table()` for counting how many tokens belong to each type


