---
title: 'Introduction to Outliers'
author: "STA 325: Lab 2, Fall 2018"
output: pdf_document
---

Today's agenda: finding outliers 

Programming partners:
You should have a programming partner for each lab, and you should switch off who is programming, and use each other for help. We will spend about 30--50 minutes per week on lab exercises and you will be expected to bring your laptops to class to work on these exercises in class. Myself and the TA will be in class to help you.

***Background***

Identifying outliers in data is an important part of statistical analyses.  One
simple rule of thumb (due to John Tukey) for finding outliers is based on the
quartiles of the data: the first quartile $Q_1$ is the value $\geq 1/4$ of the
data, the second quartile $Q_2$ or the median is the value $\geq 1/2$ of the
data, and the third quartile $Q_3$ is the value $\geq 3/4$ of the data.  The
interquartile range, $IQR$, is $Q_3 - Q_1$.  

Tukey's rule says that the outliers are values more than $1.5$ times the interquartile range from the quartiles --- either below $Q_1 - 1.5 IQR$, or above $Q_3 + 1.5 IQR$.

In this lab, we will consider the following data

```{r}
x <- c(2.2, 7.8, -4.4, 0.0, -1.2, 3.9, 4.9, 2.0, -5.7, -7.9, -4.9,  28.7,  4.9)
```

We will use these as part of writing a function to identify outliers according
to Tukey's rule.  Our function will be called \texttt{tukey.outlier}, and will
take in a data vector, and return a Boolean vector, \texttt{TRUE} for the
outlier observations and \texttt{FALSE} elsewhere.

***Lab Tasks***


1. (5) Calculate the first quartile, the third quartile, and the
  inter-quartile range of \texttt{x}.  Some built-in R functions calculate
  these; you cannot use them, but you could use other functions, like
  \texttt{sort} and \texttt{quantile}.
  
2. (10) Write a function, \texttt{quartiles}, which takes a data vector and
  returns a vector of three components, the first quartile, the third quartile,
  and the inter-quartile range.  Show that it gives the right answers on
  \texttt{x}.  (You do not have to write a formal test for \texttt{quartiles}.)

3. (5) Which points in \texttt{x} are outliers, according to Tukey's rule, if
  any?


4. (20) Write \texttt{tukey.outlier}, using your \texttt{quartiles}
  function.  The function should take a single data vector, and return a
  Boolean vector, take in a data vector, and return a Boolean vector,
  \texttt{TRUE} for the outlier observations and \texttt{FALSE} elsewhere.
  Show that it passes \texttt{test.tukey.outlier}.
  
5. (20) Write a function, \texttt{test.tukey.outlier}, which tests the
  function \texttt{tukey.outlier} against your answer in the previous question.
  This function should return \texttt{TRUE} if \texttt{tukey.outlier} works
  properly; otherwise, it can either return \texttt{FALSE}, or an error
  message, as you prefer. 
  
6. (5) Which data values should be outliers in \texttt{-x}?


7. (5) Which data values should be outliers in \texttt{100*x}?

8. (10) Modify \texttt{test.tukey.outlier} to include tests for these cases.

9. (5) Show that your \texttt{tukey.outlier} function passes the new set of
  tests, or modify it until it does.

10. (15) According to Tukey's rule, which points in the next vector $y$ are
  outliers?  What is the output of your function?  If they differ, explain why.

```{r}
y <- c(11.0, 14.0, 3.5, 52.5, 21.5, 12.7, 16.7, 11.7, 10.8, -9.2, 12.3, 13.8, 11.1)
```





